---
title: "Assignment 6"
author: "Glenn Niles"
date: "10/13/18"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Factors affecting the price of used cars

Load necessary libraries here.
```{r} 
library(dplyr)
library(ggformula)
library(readr)
```

a.	Download Cars 2005.csv and load the data into R.  Display the first few lines of data.

(Dataset:  "Car Data," submitted by Shonda Kuiper, Grinnell College. Dataset obtained from the Journal of Statistics Education (http://www.amstat.org/publications/jse). Accessed 3 June 2015. Used by permission of author.)


```{r}
cars = read_csv("Cars 2005.csv")
head(cars)
```

***Note:**  All of parts b-k should be done using functions from `dplyr` and/or `ggformula`.

b. The variables Cruise, Sound, and Leather use 1 to represent "yes" (the car has cruise control, an upgraded sound system, or a leather interior) and 0 to represent "no".  Therefore, these variables are categorical--but R treats them as numeric, because they look like 0's and 1's.

The function `as.factor()` will return a categorical variable version of its argument (the variable you put inside the parentheses).  Use `mutate` and `as.factor` to convert the variables Cruise, Sound, and Leather into factor variables.

```{r}
cars = cars %>%
        mutate(Cruise = as.factor(Cruise), Sound = as.factor(Sound), Leather = as.factor(Leather))
```

c. Make a scatterplot displaying the relationship between Price and Liter (a measure of engine size).  

```{r}
gf_point(Price ~ Liter, data = cars)%>%
  gf_labs(title = "Price as a function of Liter",
          x = "Liter",
          y = "Price",
          caption = "Source: Cars 2005")
```

- **Write a sentence** explaining which variable you chose as `x`, and why.
I chose my Liter to be my x-axis. At first, I had price/Liter and the graph looked quite messy. After switching the plot looks cleaner and its easier to see the relationship
between the two variables.
- **Write a sentence** describing the relationship between these two variables.
The relationship between price and liter isn't as strong as I predicted it would be. It still seems like in general that the bigger the engine the more the car costs. I think the 4.5L engines are throwing me off but those could be the expensive German cars too. 
So in general, the larger the engine the more expensive the car.

d. Modify your scatterplot to include the variable Leather, in addition to Price and Liter. **Include** an informative title for your graph.  

```{r}
gf_point(Price ~ Liter, col =~ Leather, data = cars)%>%
  gf_refine(scale_color_manual(values = c("red", "green")))%>%
    gf_labs(title = "Price as a function of Liter",
            x = "Liter",
            y = "Price",
            caption = "Source: Cars 2005")
```


- **Write a sentence** explaining which attribute you mapped to Leather, and why.
I used color to map the Leather variable. Since there is only two possibilities for Leather it seemed to make sense to use one color to represent one option
and a second color to represent the other.

e. Make a histogram of Price.  Use vertically-stacked facets to display the price for cars with and without leather interiors.

```{r}
gf_histogram(~ Price, data = cars) %>%
  gf_facet_grid(Leather ~ .)
```

f. Make a histogram of Price, with side-by-side bars for cars with and without leather.

```{r}
gf_histogram(~ Price, data = cars) %>%
  gf_facet_grid(. ~ Leather)
```

g. Make *one* barplot showing how many cars there are of each Type (Sedan, Convertible, etc.) and Leather value.

```{r}
gf_bar(~ Type, fill =~ Type, position = position_dodge(), data = cars)
```


h. Create a new variable in the `cars` data frame, called Tier, which classifies cars into 3 categories:  Luxury, Standard, and Economy.  Your categorization may be based on any variables in the `cars` data frame *except* Price.

```{r}
cars <-
  cars %>%
  mutate(Tier = case_when(
    Make == "Cadillac" | Cylinder == 8  ~ "Luxury",
    (Make == "Chevrolet" & Cylinder == 6) | Make == "SAAB" | Type == "Convertible" ~ "Standard",
    TRUE ~ "Economy"
  ))
```


i.  Make a dataframe showing the median price of cars of each Tier.  Display the results in increasing order of price.

```{r}
cars %>%
  group_by(Tier) %>%
  summarise(median = median(Price))
```


- **Write a sentence** discussing whether your classification makes sense:  Is the median price higher for Luxury cars than for Standard cars?  Is the median price higher for Standard cars than for Economy cars?

After a lot of messing around I think I got it so its pretty even and the prices match what I would consider fair prices per Tier while having variables that make sense. Cadillacs and V8s aren't going to be economy cars, Pontiac Sunfires wouldn't be luxury, even if you got a cool convertible one, etc.

j. Make a barplot showing the median price of cars of each Tier.
- Optional:  Use code to ensure that the results will be displayed in increasing order of price.

```{r}
medianTier <-
  cars %>%
  group_by(Tier) %>%
  summarise(medianPrice = median(Price))
medianTier%>%
mutate(Tier = reorder(Tier, medianPrice)) %>%
gf_col(medianPrice ~ Tier)
```

k. Optional:  Use `select` and `left_join` to produce a data frame containing the columns Price, Mileage, Tier, and the median price for cars of that tier.  Display the first few rows of the data frame.

```{r}
carsEC <- select(cars, Price, Mileage, Tier)
carsEC <- left_join(carsEC, medianTier)
head(carsEC)
```

